Hierarchical time series forecasting in Emergency Medical Services

### 2022-11-16 manuscript/main.qmd
Updated abstract

---
title: "Hierarchical time series forecasting in Emergency Medical Services"
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  The primary goal of an emergency service is to minimize the ambulance response time. Accurate forecasts of ambulance demand are crucial inputs into the decisions of planners and policymakers on the resource (e.g. staff and fleet) planning and dynamic deployment. Informing those decisions require detailed forecasting of demand at multiple levels of hierarchy from the national and regional levels to the nature of incidents and their priorities. These forecasts are often generated independently by different teams. However, independent forecasts may not add up to the other levels resulting in mismatched forecasts that lead to conflicting decisions and a lack of coordination in the care service. To address this issue, we explore the inherently hierarchical and grouped structure of Emergency Medcine Services. We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
keywords: 
  - forecasting
  - healthcare
  - emergency services
  - forecast reconciliation
  - hierarchical time series
  - ambulance demand
  - atteneded incidents

>>>>

---
title: "Hierarchical time series forecasting in Emergency Medical Services"
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet. Such demand forecasts are required at national, regional and sub-regional levels, and must take account of the nature of incidents and their priorities. These forecasts are often generated independently by different teams within the organization. As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service. To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation. The methods are applied to daily incident data from the Welsh Ambulance Service NHS Trust, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area. We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model. We validate the forecasting approach using time-series cross-validation.
keywords:
  - forecasting
  - healthcare
  - emergency services
  - forecast reconciliation
  - hierarchical time series
  - ambulance demand
  - attended incidents
